---
title: "Mixed Model"
author: "C. M. Ciafre"
date: "September 17, 2019"
output: html_document
---
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages_needed <- c("ggplot2", # graphics
                     "dplyr",
                     "arm", # display() etc.
                     "MASS",
                     "ggfortify",
                     "nlme",
                     "lme4",
                     "recluster",
                     "cluster")
pk_to_install <- packages_needed [!( packages_needed %in% rownames(installed.packages())  )]
if(length(pk_to_install)>0 ){
  install.packages(pk_to_install,repos="http://cran.r-project.org")
}
#lapply(packages_needed, require, character.only = TRUE)
library(ggplot2)
library(dplyr)
library(arm)
library(MASS)
library(ggfortify)
library(nlme)
library(lme4)
library(recluster)
library(cluster)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

vegdata<-read.csv("data/VEGDATADONE.csv", header=TRUE)
abbrevcols<-read.csv("Data/AbbrevCols.Fall2019.csv", header=TRUE)
```
```{r Prepare the data, include=FALSE}
quadrich1<-count(vegdata, ID)
quadrich<- quadrich1[-c(1),]
incompdata<-cbind(quadrich, abbrevcols)
```
### Background

I collected plant species data from 24 wetland sites across the Eastern Highland Rim ecoregion once in early summer and once in late summer. Each site was subsampled with a varying number of 1 m^2 quadrats, resulting in 923 subsamples.

Fixed effects:
  -Sampling season (early or late)
  -Water permanence status (ephemeral or permanent)
  -Site size class (small or medium or large)
  
Random effects:
  -Plot
  -Quadrat

### Two ways to visualize trends in the data

We can visualize all the quadrats at once and see if sampling season and water permanence affect quadrat level richness...
(Note that sampling effort was halved for the late season to save time)
```{r echo=FALSE}
ggplot(incompdata, aes(x=Season, y=n, group=Permanence, color=Permanence))+
  geom_line()+
  geom_point()+
  geom_jitter(width = 0.3)+
  xlab("Sampling Season")+
  ylab("Quadrat Richness")+
  ggtitle("Quadrat richness between seasons")
```

Or we can examine the effects of season on each site using mean quadrat richness...
```{r include=FALSE}
meanbyplot <- aggregate(n ~ Pond:Season:Permanence:MTLength, data=merged, mean)
```
```{r echo=FALSE}
ggplot(meanbyplot, aes(x=Season, y=n, color=Permanence))+
  geom_line(aes(group=Pond))+
  geom_point()+
  xlab("Sampling Season")+
  ylab("Mean Quadrat Richness")+
  ggtitle("Mean quadrat richness of plots across seasons")
```

It also appears that site size has an effect on mean quadrat richness...
(I used size classes rather than site radius to make things simpler)
```{r include=FALSE}
meanbyplot3 <- aggregate(n ~ Pond:Permanence:Size, data=merged, mean)
```
```{r echo=FALSE}
ggplot(meanbyplot3, aes(x=Size, y=n, color=Permanence))+
  geom_point()+
  xlab("Site Radius (m)")+
  ylab("Mean Quadrat Richness")+
  ggtitle("Mean quadrat richness among sites of different sizes")
```

### Let's make some models
```{r message=FALSE, warning=FALSE}
#Fully crossed
model1<-lmer(n~Permanence+Size+Season+Permanence:Size+Size:Season+Season:Permanence+Permanence:Size:Season+(1|Pond:Quadrat), data= incompdata)

#Permanence:Size interaction only
model2<-lmer(n~Permanence+Size+Season+Permanence:Size+(1|Pond:Quadrat), data= incompdata)

#No interactions
model3<-lmer(n~Permanence+Size+Season+(1|Pond:Quadrat), data= incompdata)

#model2 without Season
model4<-lmer(n~Permanence+Size+Permanence:Size+(1|Pond:Quadrat), data= incompdata)
```

```{r}
anova(model1, model2)
```

Model 2 did not significantly differ from model 1 (p=0.632), so it is likely that the discarded interactions (the ones that did not make biological sense) were unimportant to the model.

```{r}
anova(model1, model3)
anova(model2, model3)
```

To see if the remaining interaction term was meaningful, I removed it from model 3; as a consequence, model 3 was significantly different from both 1 and 2 (p=1.46e-4 and 2.71e-05). This indicates that the interaction between permanence and size is important to the model and should be left in.

```{r}
anova(model2, model4)
```

Because I was particularly interested in whether quadrat richness changed between sampling season, I removed that term from model 2. Because model 4 significantly differed from model 2 (p=0.00792), sampling season is likely important to the model and should be kept in. 

